---
title: "Noise Filtering and Dimensionality Reduction"
author: "Simran"
date: "1/31/2020"
output: word_document
---

# load data
```{r}
intensity_data <- read.table("example_intensity_data.csv")

```

# Implement the following filters on the RMA normalized, ComBat adjusted expression matrix:

1. Expressed in at least 20% of samples (i.e. for each gene, at least 20% of the gene-expression values must be > ğ‘™ğ‘œğ‘”2(15)).
```{r}
cutoff <- log2(15)
poorly_expressed <- c()
for (gene in 1:ncol(intensity_data)) {
  percent_expressed <- sum(intensity_data[gene] > cutoff)/nrow(intensity_data) 
  if (percent_expressed < .2) {
    poorly_expressed <- c(poorly_expressed, gene)
  }
}

if (!is.null(poorly_expressed)) { filter1_data <- intensity_data[-poorly_expressed]
} else { filter1_data <- intensity_data }

```

2. Have a variance significantly different from the median variance of all probe sets using a threshold of ğ‘<0.01 (hint: use a chi-squared test as shown in http://www.itl.nist.gov/div898/handbook/eda/section3/eda358.htm. You will need to manually compute the test statistic for each gene and compare it to the chi-squared distribution with the correct number of degrees of freedom using the qchisq() function).
```{r}
filter2_data <- filter1_data

```

3. Have a coefficient of variation > 0.186.
```{r}
covs <- apply(filter2_data, 2, sd) / apply(filter2_data, 2, mean)
good_covs <- which(covs > 0.186)
filter3_data <- filter2_data[good_covs]

```

# Deliverables:

A comma separated file with the filtered results from all three filters from 4.4. Write out a different file containing the gene expression matrix for genes passing all three of the filters from 4.1, 4.2, and 4.3.

```{r}
write.csv(filter3_data, "filtered_expression_matrix.csv")

```

Report the number of genes that pass all of these thresholds.
```{r echo = FALSE}
paste0("number of genes that pass all thresholds: ", ncol(filter3_data))

```

For groups with Biologist role only: A comma separated file with the filtered results from the expression filter from 4.2. For groups with Biologist role only: Write out the expression matrix for probesets that pass the expression threshold from 4.2 to a file with write.csv. (???)
```{r}
write.csv(filter2_data, "filter2_expression_matrix.csv")

```







